---
title: "Diagnostics for GP emulators"
author: "Victoria Volodina"
date: "06/05/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height=3, fig.width=10)
```

This vignette is aiming to demonstrate how to perform diagnostics for emulator object obtained with ExeterUQ MOGP. 

### Preliminaries

First we specify the directory where mogp is installed so that the python is correctly imported. Your directory will be different from mine.

```{r}
mogp_dir <- "/Users/vvolodina/Documents/mogp_emulator"
```

```{r results='hide', message=FALSE, warning=FALSE}
setwd('..')
source('BuildEmulator/BuildEmulator.R')
```

The description of data format and how to build mogp emulators are considered in detail in other vignettes.

We start by loading a `.Rdata` file, which contains a data frame object `tData`. The first three columns correspond to the input parameters, while the last three columns after the `Noise` correspond to the model outputs (metrics of interest).
We retain the first 25 points as our training set, and the last 5 points will be used for validation.

```{r}
load("ConvectionModelExample.Rdata")
names(tData)
cands <- names(tData)[1:3]
print(cands)
tData.train <- tData[1:25, ]
tData.valid <- tData[26:30, ]
```

We proceed to construct a GP emulator with default settings for all three model outputs.

```{r results='hide', message=FALSE, warning=FALSE}
TestEm <- BuildNewEmulators(tData = tData.train, HowManyEmulators = 3)
```

### Leave-One-Out (LOO) diagnostics

After generating a GP object, we perform Leave-One-Out (LOO) diagnostics. We specify our obtained GP object as `Emulators` argument of `LOO.plot` function. We define the index of the emulator for which we want to produce LOO diagnostics plot `which.emulator`, i.e. we are interested to produce diagnostics for the first emulator. `ParamNames` is a vector of names of input parameters.

```{r}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 1,
                  ParamNames = cands)
print(head(tLOOs))
```


In the LOO diagnostics plot, the black dots and error bars show predictions together with two standard deviation prediction intervals, while the green/red points are the true model output coloured by whether or not the truth lies within the error bars.

By calling `LOO.plot` function, we also obtain a data frame with three columns, with first column corresponding to posterior mean, and second and third columns corresponding to the minus and plus two standard deviations.

We can change the order and/or the number of input parameters in `ParamNames` specification.

```{r}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 1,
                  ParamNames = c("A_T", "A_U", "A_EPSILON"))
```

```{r, fig.height=3, fig.width=7}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 1,
                  ParamNames = c("A_T", "A_EPSILON"))
```

Notice that by choosing to produce LOO plot for the third emulator, we only have two input variables in our LOO diagnostics plots, since only these two input variables are active.
```{r}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 3,
                  ParamNames = cands)
head(TestEm$fitting$Design[, TestEm$fitting$ActiveIndices[[3]]])
```

#### LOO plots on original scale
Modellers are interested in studying these plots on the original parameter scale. With `LOO.plot` function, they have an option to specify `OriginalRange=TRUE`. Those ranges are read from a file containing the parameters ranges and whether the parameters are logged or not. The string that corresponds to the name of this file is provided in `RangeFile`.


```{r}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 3,
                  ParamNames = names(TestEm$fitting.elements$Design), 
                  OriginalRanges = TRUE, RangeFile="ModelParam.R")
```

#### LOO plots with Observation and Observation Error

Modellers who are aiming to perform history matching with our emulators could be interested in adding the information about the observation, `Obs`, and the observation error, `ObsErr`. 

```{r}
tLOOs <- LOO.plot(Emulators = TestEm, which.emulator = 3,
                  ParamNames = names(TestEm$fitting.elements$Design), 
                  OriginalRanges = TRUE, RangeFile="ModelParam.R", 
                  Obs = 14, ObsErr = 0.1)
```
We specified observation value at 14 together with observation error at 0.1. The blue dashed lines correspond to the observation together with plus/minus two observation error.


## Validation plots

A stener validation test is to analyse the emulator performance on unseen data set, i.e. `tData.valid`.

```{r}
tValid <- ValidationMOGP(NewData = tData.valid, 
                         Emulators = TestEm, which.emulator=3,
                         tData = tData, ParamNames = cands)
print(head(tValid))
```

Similar to LOO diagnostics plots, the black points and error bars show predictions together with two standard deviation prediction intervals, while the green/red points are the true model output coloured by whether or not the truth lies within the error bars.

By calling `ValidationMOGP` function, we also obtain a data frame with three columns, with first column corresponding to posterior mean, and second and third columns corresponding to the minus and plus two standard deviations

In case we previously used `ValidationMOGP` function and saved the prediction object, we can re-use this function to produce the validation plots. We specify `tValid` inside the `Predictions` of the function.

```{r}
ValidationMOGP(NewData = tData.valid, 
               Emulators = TestEm, which.emulator=3,
               tData = tData, ParamNames = cands, 
               Predictions = tValid)

```

### Validation plots on original scale

We have an option to produce plots on original input parameter scales by specifying `OriginalRanges=TRUE`. Those ranges are read from a file containing the parameters ranges and whether the parameters are logged or not. The string that correspond to the name of this file is provided in `RangeFile`.

```{r}
tValid <- ValidationMOGP(NewData = tData.valid, 
                         Emulators = TestEm, which.emulator=3,
                         tData = tData, ParamNames = cands, 
                         OriginalRanges = TRUE, 
                         RangeFile= "ModelParam.R")
```

### Validation plots with Observation and Observation Error

Modellers have an option to add information about the observation and observation error in their plots. We specified observation value, `Obs`=15, and observation error, `ObsErr`=0.1. The blue dahsed lines correspond to the observation value plus/minus two observation error value

```{r}
tValid <- ValidationMOGP(NewData = tData.valid, 
                         Emulators = TestEm, which.emulator=3,
                         tData = tData, ParamNames = cands, 
                         OriginalRanges = TRUE, 
                         RangeFile= "ModelParam.R", 
                         Obs=14, ObsErr=0.1)
```
