---
title: "Saving ExeterUQ_mogp emualtors"
author: "Danny Williamson"
date: "15/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Saving and loading

```ExeterUQ_mogp``` emulators are not standard objects. They are lists with 2 elements: the first of which is a python mogp_emulator object, and the second is a list of statistical elements of the fit such as prior choices, mean function choices, elements useful for diagnostics and other things we would like for transparent inference.

The usual ```save()``` and ```load()``` functions will seem to work, but will only save a pointer for the python object (which once removed/reloaded won't work). Here is a MWE

```{r}
mogp_dir <- "~/Dropbox/BayesExeter/mogp_emulator"
```

```{r results='hide', message=FALSE, warning=FALSE}
setwd('..')
source('BuildEmulator/BuildEmulator.R')
```

```{r}
load("ConvectionModelExample.Rdata")
TestEm <- BuildNewEmulators(tData, HowManyEmulators = 2)
print(TestEm[[1]])
save(TestEm, file="TestEm.RData")
rm(TestEm)
load("TestEm.RData")
print(TestEm[[1]])
```

The mogp part has gone on reload (it was never saved). To overcome this we use the python functions ```py_save_object``` and ```py_load_object``` to save out the mogp part. This means that our saving and loading of an mogp emulator actually saves an RData file and a python object separately. This is all handled automatically within our package using the functions ```saveExUQmogp``` and ```loadExUQmogp``` 
First we build an emulator again
```{r}
TestEm <- BuildNewEmulators(tData, HowManyEmulators = 2)
```

Now to save we call
```{r}
save_ExUQmogp(TestEm, filename = "SavedEmulator")
```

It is important that the filename here has no extensions. Within your directory, this function has now saved 2 files: ```SavedEmulator.RData``` is the Rlist and ```SavedEmulator_mogp``` is the Python object. 

To load an emulator, both the R and Python objects have to exist in the directory (so you would be loading an emulator that was saved using ```save_ExUQmogp```). Removing ```TestEm``` to check the call we have

```{r}
rm(TestEm)
TestEm <- load_ExUQmogp("SavedEmulator")
```

Checking successful loading:

```{r}
newDesign <- 2*randomLHS(100,3)-1
preds <- TestEm$mogp$predict(newDesign, deriv=FALSE)
preds$mean
```
